---
title: "Waze Leaflet Exploration"
output: 
  md_document:
    variant: markdown_github
always_allow_html: yes
---

### Import the relevant libraries and odbc connectors.

```{r}
library(RODBC)
library(tidyverse)
library(leaflet)
library(rgdal)
library(lubridate)
library(stringr)
source('/home/michelle/michelle.tat@boston.gov/Work Projects/Waze Dashboard/dashboard/data_wrangling.R')
source('/home/michelle/michelle.tat@boston.gov/Work Projects/Waze Dashboard/dashboard/aggregation.R')
source('/home/michelle/michelle.tat@boston.gov/Work Projects/Waze Dashboard/dashboard/plotting.R')
```

### Create a database connection.
```{r}
conn <- odbcDriverConnect(conn_string)

```

### Get a list of streets and end nodes.
```{r}
jam_streets <- read_csv('/home/michelle/michelle.tat@boston.gov/Work Projects/Waze Dashboard/dashboard/data/jam_streets.csv')
```

### Define street names
```{r}
street_name <- 'Washington St'
an_end_node <- 'South St'
```

### Define month, day and year.
```{r}
selected_month <- 'Oct'
selected_day <- 20
selected_year <- 2017
```


### Create a query to retrieve 
```{r}
custom_query <- sprintf("SELECT DISTINCT jam.city,
jam.street,
jam.endNode,
jam.speed,
jam.delay,
jam.startTime AS startTime,
jam.endTime AS endTime,
jam.uuid
FROM Waze.dbo.Waze_JamData AS jam
WHERE city = 'Boston, MA' and street = '%s' and jam.endNode = '%s' and jam.startTime >= '2017-01-01 00:00:00' and jam.startTime <= '2017-12-31 00:00:00'
ORDER BY startTime ASC", street_name, an_end_node)
```

### Run the query
```{r}
waze_query <- sqlQuery(conn, custom_query)
```

### Get x-y coordinate to plot later
```{r}
street_coordinate <- sqlQuery(conn, sprintf("SELECT TOP 30 jam.street, 
                              jam.endNode, 
                              point.x, 
                              point.y
                              FROM dbo.Waze_JamData AS jam
                              INNER JOIN dbo.Waze_PointData AS point
                              ON jam.uuid=point.uuid
                              WHERE jam.street='%s' and jam.endNode = '%s'", street_name, an_end_node))
```
  
### Dedupe logic:
### uuids are indicators of a jam, usually last < 1hr
### lots of empty uuids, so we have to figure out the nature of those entries
### group entries that are approx an hour apart
### assign uuids to those entries 
### re-add them to origin df
```{r}
city <- as.vector(waze_query$city[1])
street <- as.vector(waze_query$street[1])
endNode <- as.vector(waze_query$endNode[1])

empty_uuids <- waze_query %>% filter(uuid == '') %>% mutate(time_diff = endTime - startTime, 
                                                                     month_startTime = month(startTime), 
                                                                     day_startTime = day(startTime),
                                                                     hour_startTime = hour(startTime),
                                                                     minute_startTime = minute(startTime),
                                                                     month_endTime = month(endTime), 
                                                                     day_endTime = day(endTime),
                                                                     hour_endTime = hour(endTime),
                                                                     minute_endTime = minute(endTime),
                                                                     start_end_diff = hour_endTime - hour_startTime)

empty_uuids <- empty_uuids %>% group_by(day_startTime, hour_startTime) %>% 
                              summarise(count = n(),
                              speed = mean(speed, na.rm=TRUE),
                              delay = mean(delay, na.rm=TRUE),
                              startTime = min(startTime),
                              endTime = max(endTime)) %>%
                              mutate(n_jams = 1,
                                     city = city,
                                     street=street,
                                     endNode=endNode)
                  

empty_uuids$uuid <- sample(100000000000, size = nrow(empty_uuids), replace=TRUE)

empty_uuids <- empty_uuids %>% ungroup(day_startTime) %>% select(-day_startTime)


empty_uuids <- empty_uuids %>% ungroup(day_startTime, hour_startTime) %>% 
  select(city, street, endNode, uuid, count, n_jams, speed, delay, startTime, endTime) %>%
  mutate(uuid = as.character(uuid))

```

```{r}
examine_dedupe <- waze_query %>% 
                  filter(uuid != '') %>%
                  group_by(city, street, endNode, uuid) %>% 
                  summarise(count = n(),
                            n_jams = 1,
                            speed = mean(speed, na.rm=TRUE),
                            delay = mean(delay, na.rm=TRUE),
                            startTime = min(startTime), 
                            endTime = max(endTime)) %>%
                  mutate(uuid = as.character(uuid))
```  

```{r}
waze_deduped <- examine_dedupe %>% bind_rows(empty_uuids)
```


```{r}
  waze_deduped <- waze_deduped %>%
    mutate(startTime=ymd_hms(startTime, tz = 'EDT')) %>%
    mutate(month=month(startTime, label = TRUE)) %>%
    mutate(day=day(startTime)) %>%
    mutate(weekday = wday(startTime, label = TRUE)) %>%
    mutate(hour=hour(startTime)) %>%
    mutate(year=year(startTime)) %>%
    mutate(week=week(startTime))
```


### Dedupe the waze data
```{r}
waze_deduped <- dedupe_and_format_waze(waze_query)
```

### Subset waze data to selected month and day.
```{r}
waze_df <- to_month_and_day(waze_deduped, selected_month, selected_day)

```

### Get daily information for a specific day. This creates a df grouped by hour.
```{r}

waze_daily <- aggregate_daily_metrics(waze_df)

```

### Aggregate daily information, as well as aggregate daily information over the year (e.g., all Mondays).
### Spits out a dataframe with both pieces of information.
```{r}
waze_daily_avg <- aggregate_daily_average(waze_deduped, waze_daily, selected_month, selected_day)
```

### Gets all information for the day of interest, and all n number of days back (e.g., 7 days, 14 days, 21 days)
```{r}
the_year = 2017
the_day = '30'
the_month = 'Nov'
the_date <- paste0(the_year, '-', the_month, '-', the_day)
the_date <- as.character(ymd(the_date))
waze_monthly <- aggregate_month_metrics(waze_deduped, origin_date = the_date, 21)
```


### Creates a barplot of specific waze metrics
```{r}
barplot_waze_metrics(waze_daily, 
                     waze_daily$hour_of_day, 
                     waze_daily$n_waze_reported_jams,  
                     'Hour of Day', 
                     'Number of Reported Jams')
```
### Create a barplot of daily vs annual jams
```{r}
barplot_waze_metrics(waze_daily_avg,
                     waze_daily_avg$category,
                     waze_daily_avg$mean_total_jams,
                     'Daily or Annual',
                     'Number of Jams')
```

### Create a lineplot of historical waze info
```{r}
lineplot_waze_metrics(waze_monthly,
                      waze_monthly$date_label,
                      waze_monthly$n_waze_reported_jams,
                      "Date",
                      "Number of Reported Jams")
```


### Create a leaflet map indicate the corridor of interest.
```{r}
leaflet(street_coordinate) %>% addPolylines(lng=~x, lat=~y, opacity = 1, noClip=TRUE, stroke=1) %>% addProviderTiles(providers$Stamen.Toner)
```

